---
title: "World Value Survey - Exploration"
author: "Nicolas Restrepo"
date: "2023-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(countrycode)
library(janitor)
library(cmdstanr)
theme_set(theme_bw())
```

I am going to use this document to explore the latest time series from the World Values Survey. 

Let's begin by loading in the data:

```{r}
load("Data/wvs_timeseries.rdata")
d <- data1
rm(data1)

head(glimpse(d))
```

From looking at the codebook, I think there are seven questions we can focus on. 

1. Whether homosexuality is ever justifiable (nine-point scale from `never justifiable` to `always justifiable`).

2. Whether prostitution is ever justifiable (nine-point scale from `never justifiable` to `always justifiable`).

3. Whether abortion is ever justifiable (nine-point scale from `never justifiable` to `always justifiable`).

4. Whether divorce is ever justifiable (nine-point scale from `never justifiable` to `always justifiable`).

5. Whether euthanasia is ever justifiable (nine-point scale from `never justifiable` to `always justifiable`).

6. Whether suicide is ever justifiable (nine-point scale from `never justifiable` to `always justifiable`).

7. Whether you would **not** like immigrants/foreign workers as neighbors.

I am now going to select these variables and give them cleaner names, so they are easier to manipulate: 

```{r}
# Read in the data 
d_short <- d %>% 
  select(S002VS, 
         COUNTRY_ALPHA,
         S020,
         S003,
         X001,
         X003, 
         X007, 
         A001:A006,
         A029:A042,
         A124_02, 
         A124_03, 
         A124_06:A124_09,
         A165,
         A170,
         A173,
         C001,
         C002,
         D057,
         E033,
         E035:E037,
         E039,
         E069_01:E069_02,
         E069_04:E069_08,
         E069_10:E069_13,
         E069_17,
         F028,
         F034,
         F050,
         F053,
         F063,
         F114A,
         F115:F123,
         G006) %>% 
  rename(wave = S002VS, 
         country = COUNTRY_ALPHA,
         country_code = S003, 
         year_survey = S020,
         sex = X001,
         age = X003, 
         
         # Importance battery: 1 means very important, 4 means not at all important
         important_family = A001,
         important_friends = A002, 
         important_leisure = A003, 
         important_poltics = A004,
         important_work = A005, 
         important_religion = A006, 
         
         # Important qualities in children: 1 means mentioned, 0 means not mentioned
         child_independence = A029, 
         child_hard_work = A030, 
         child_feeling_responsibility = A032, 
         child_imagination = A034, 
         child_tolerance = A035, 
         child_thrift = A038, 
         child_determination = A039, 
         child_religion = A040, 
         child_unselfish = A041,
         child_obedience = A042,
         
         # People you would not like to have as neighbours: 1 mentioned, 0 means not mentioned
         neigh_diff_race = A124_02, 
         neigh_drink = A124_03, 
         neigh_imm = A124_06, 
         neigh_aids = A124_07, 
         neigh_drugs = A124_08, 
         neigh_gay = A124_09,
         
         # Most people can be trusted: (1) Most people, (2) you can never be too careful
         trust_people = A165,
         
         # (10) means more satisfied
         life_satisf = A170, 
         
         # Control over life's choices: (10) feels most in control. 
         choice_control = A173, 
         
         # Men should be given jobs over women: (1) Agree, (2) Disagree, (3) Neither A nor D
         jobs_men_over_women = C001, 
         
         # Nationals should be given jobs over foreigners: (1) Agree, (2) Disagree, (3) Neither A nor D
         jobs_national_over_foreign = C002, 
         
         # Likert: (1) Strongly Agree -- (4) Strongly Disagree
         housewife_fulfilling = D057,
         
         # Left/Right placement: (1) Left -- (10) Right
         politics_scale = E033,
         
         # Importance of income equality 
         income_eq = E035, 
         
         # Business should be public/private owned: (1) Private -- (10) Public
         pvt_state_owned = E036, 
         
         # Government should take responsibility vs individual responsibility: (1) Gvt -- (10) Individual
         gvt_responsibility = E037,
         
         # Competition good or harmful: (1) Good --- (10) harmful
         competition_good_evil = E039, 
         
         # Confidence battery: (1) A great deal; (2) Quite a lot; (3) Not very much; (4) None at all
         confidence_churches = E069_01, 
         confidence_armed_forces = E069_02,
         confidence_press = E069_04, 
         confidence_unions = E069_05, 
         confidence_police = E069_06, 
         confidence_parliament = E069_07,
         confidence_civil = E069_08,
         confidence_television = E069_10,
         confidence_governement = E069_11,
         confidence_political_party = E069_12,
         confidence_major_companies = E069_13,
         confidence_justice_courts = E069_17,
         
         # How often attends religious services?
         # (1) More than once a week -- (8) Never
         attend_relig = F028, 
         
         # (1) Religious; (2) not religious; (3) atheist
         religious_person = F034,
         
         # Believes (1), does not believe (0)
         believe_god = F050, 
         believe_hell = F053, 
         
         # (1) Not at all important -- (10) very important
         important_god = F063,
         just_gvt_benefits = F114A,
         just_fare_public_trans = F115,
         just_cheat_taxes = F116,
         just_bribe = F117,
         just_homosexuality = F118, 
         just_prostitution = F119, 
         just_abortion = F120, 
         just_divorce = F121, 
         just_euthanasia = F122, 
         just_suicide = F123, 
         proud_nationality = G006, 
         marital_status = X007) 



# Zap labels 
d_short <- haven::zap_labels(d_short)
  
```

Our data  is organized in such a way where each wave is an individual, in a country, in a given wave. The first thing to look at is how available our questions of interest are in each country and each wave. Before we do that, we need to deal with the missing values a bit. 

```{r}
# All missing values are indicated with negative values
# Turn them into NAs for now

d_short <- d_short %>% 
  mutate(
    across( 
      .cols = 5:68, 
       ~ ifelse(
        . < 0, 
        NA_real_, 
        .
      ))
  )
```

I am going to remove the `NAs` and examine how many people I have per country per wave. 

```{r}


summary_df <- d_short %>% 
  select(country,
         wave,
         8:14) %>% 
  pivot_longer(3:9, 
               names_to = "variable", 
               values_to = "response") %>% 
  mutate(valid_response = if_else(
    is.na(response),0, 1
  )) %>% 
  group_by(
    country, 
    wave, 
    variable
  ) %>% 
  summarise(
    count = sum(valid_response)
  ) %>% 
  ungroup()
  
summary_df %>% 
  ggplot(aes(
    x = wave, 
    y = count, 
    fill = variable
  )) +
  geom_bar(stat = 'identity') +
  facet_wrap(~country) +
  theme(axis.text.y = element_blank()) +
  theme(legend.position = 'none')
```

We have countries for which we have pretty comprehensive data like the USA, Sweden, Canada and Argentino. Other countries, like Uganda only have one wave. 

### Some Trajectories 

I am going to select countries for which we have more than four waves and look at some of the trajectories of our data.

Let's look at homosexuality:

```{r}

wave_df <- d_short %>% 
  group_by(country, wave) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(country) %>% 
  summarise(waves_counted = n())

d_short <- d_short %>% 
  left_join(., wave_df, by = "country")

d_short %>% 
  filter(waves_counted > 3) %>% 
  group_by(
    country, 
    wave
  ) %>% 
  summarize(
    mean_value = mean(just_homosexuality, 
                      na.rm = T)
  ) %>% 
  ggplot(aes(
    x = wave, 
    y = mean_value, 
    color = country
  )) +
  geom_point() +
  geom_line() +
  facet_wrap(~country) + 
  theme(legend.position = "none") +
  labs(title = "Is homosexuality ever justifiable?")
```

Look at that dip in India! 

What about the question of whether abortion is ever justifiable

```{r}
d_short %>% 
  filter(waves_counted > 3) %>% 
  group_by(
    country, 
    wave
  ) %>% 
  summarize(
    mean_value = mean(just_abortion, 
                      na.rm = T)
  ) %>% 
  ggplot(aes(
    x = wave, 
    y = mean_value, 
    color = country
  )) +
  geom_point() +
  geom_line() +
  facet_wrap(~country) + 
  theme(legend.position = "none") +
  labs(title = "Is abortion ever justifiable?")

```

Okay let's try to look more closely at different age-groups, to see if we can detect, at least visually, patterns in where changes are happening. I am going to zoom in on the USA, Sweden, Korea, and India for this. 

```{r}
# Categorical variable for age
d_short <- d_short %>% 
  mutate(age = if_else(age < 0, 
                       NA_real_, 
                       age), 
         age_banded = case_when(
           age < 25 ~ "<25", 
           age >= 25 & age < 35 ~ "25-34", 
           age >= 35 & age < 45 ~ "35-44", 
           age >= 45 & age < 55 ~ "45-54", 
           age >= 55 & age < 65 ~ "55-64", 
           age >= 65 ~ "65+"
         ))

d_short %>% 
  filter(country %in% c("USA", "SWE", "IND", "KOR"), 
         !is.na(age)) %>% 
  group_by(wave, country, age_banded) %>% 
  summarise(mv = mean(just_divorce, na.rm = T)) %>% 
  ggplot(aes(x = wave, 
             y = mv, 
             color = age_banded)) +
  geom_point() +
  geom_line(alpha = 0.5) +
  facet_wrap(~country) +
  labs(title = "How Justifiable is Divorce?", 
       subtitle = "Age groups across four countries")
```

```{r}

just_usa_jpn <-  d_short %>% 
  filter(country %in% c("USA", "JPN"), 
         !is.na(just_homosexuality)) %>% 
  mutate(outcome = scale(just_homosexuality, center = T, scale = T)[,1], 
         wavef = as.factor(wave))

pred_matrix <- model.matrix(~ wavef, 
                  data=just_usa_jpn)

stan_data <- list(
      x = pred_matrix, 
      outcome = just_usa_jpn$outcome, 
      N_obs = nrow(just_usa_jpn), 
      K = ncol(pred_matrix), 
      cid = ifelse(just_usa_jpn$country=="USA", 1, 2), 
      N_cts = length(unique(just_usa_jpn$country)))

simple_usa <- cmdstan_model(stan_file = "simple_mu_usa.stan")

simple_usa_fit <- simple_usa$sample(
  data = stan_data, 
  chains = 4, 
  iter_warmup =100, 
  iter_sampling =  400,
  refresh = 50, 
  parallel_chains = 4
)

draws_df <- simple_usa_fit$draws(
  format = "df"
)

expected_means <- tibble(
  wave_1 = draws_df$`beta_c[1,1]`,
  wave_2 = draws_df$`beta_c[1,1]` + draws_df$`beta_c[1,2]`,
  wave_3 = draws_df$`beta_c[1,1]` + draws_df$`beta_c[1,3]`, 
  wave_4 = draws_df$`beta_c[1,1]` + draws_df$`beta_c[1,4]`,
  wave_5 = draws_df$`beta_c[1,1]` + draws_df$`beta_c[1,5]`, 
  wave_6 = draws_df$`beta_c[1,1]` + draws_df$`beta_c[1,6]`,
  wave_7 = draws_df$`beta_c[1,1]` + draws_df$`beta_c[1,7]`
) %>% pivot_longer(
  1:7, 
  names_to = "waves",
  values_to = "estimates"
) %>% 
  mutate(
    year = case_when(
      waves == "wave_1" ~ 1, 
      waves == "wave_2" ~ 2,
      waves == "wave_3" ~ 3,
      waves == "wave_4" ~ 4,
      waves == "wave_5" ~ 5,
      waves == "wave_6" ~ 6, 
      waves == "wave_7" ~ 7
    )
  )

expected_means %>% 
  ggplot(
    aes(
      x = year, 
      y = estimates
    )
  ) +
  geom_jitter(pch = 21, 
              alpha = 0.05) +
  geom_point( 
    data = just_usa_jpn %>%
      filter(country == "USA") %>% 
    group_by(wave) %>% 
    summarise(m = mean(outcome)), 
    aes(
      x = wave, 
      y = m
    ), 
    pch = 22, 
    col = "red")
```

