---
title: "SV Exploration"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(countrycode)
library(janitor)
library(glmmTMB)
library(patchwork)
library(performance)
library(ggrepel)
theme_set(theme_bw())
```

# DATA PREP

I am going to take all of Nico's cleaning code.

```{r}
load("Data/wvs_timeseries.rdata")

d <- data1 %>% 
  haven::zap_labels()

rm(data1)

d_short <- d %>% 
  select(S002VS, 
         COUNTRY_ALPHA,
         S020,
         S003,
         X001,
         X003, 
         X007, 
         F118:F123, 
         A124_06) %>% 
  rename(wave = S002VS, 
         country = COUNTRY_ALPHA,
         country_code = S003, 
         year_survey = S020,
         sex = X001,
         age = X003, 
         marital_status = X007, 
         just_homosexuality = F118, 
         just_prostitution = F119, 
         just_abortion = F120, 
         just_divorce = F121, 
         just_euthanasia = F122, 
         just_suicide = F123, 
         neigh_imm = A124_06) 

d_short <- d_short %>% 
  mutate(
    across( 
      .cols = 6:14, # changed to include age and marital status
       ~ ifelse(
        . < 0, 
        NA_real_, 
        .
      ))
  )

```

I want to see start and end dates for individual questions in each country and the number of waves. This tells us how long the series is in that country for that question and how many distinct waves. 

NOTE: I am also going to make a few new variables:
- `year0` for models
- `age01` (makes interpretation a bit easier)
- `birthyear` (we will use more versions of this later)
- `cohort5` and `cohort10` (coarsened)

```{r}
d_waves <- d_short %>% 
  pivot_longer(just_homosexuality:neigh_imm,
               names_to = "variable",
               values_to = "y") %>%
  mutate(age01 = (age - 25)/(64 - 25),
         birthyear = year_survey - age,
         year0 = year_survey - 1981,
         cohort5 = floor(birthyear/5) * 5,
         cohort10 = floor(birthyear/10) * 10) %>%
  drop_na(c(y, birthyear)) %>% 
  group_by(country, variable) %>% 
  mutate(start = min(year_survey),
         end = max(year_survey),
         span = end - start,
         waves = n_distinct(wave),
         obs = n()) %>% 
  ungroup()
```

Keep country/variables where we have at least 4 waves over 30 years. Only use people who are 25 and above.

```{r}
d30 <- d_waves %>% 
  filter(waves >= 4 & span >= 30 & age > 24)

table(d30$country, d30$variable)
```

This might be good enough: 8 countries and in different areas of the world.

We probably need to trim the cohorts so no cell is too small.

```{r}
d30 <- d30 %>% 
  mutate(cohort10t = case_when(
    cohort10 < 1920 ~ 1920,
    cohort10 > 1980 ~ 1980,
    TRUE ~ cohort10
  ))
```


# FUNCTIONS

Estimate an RE model for each country/variable combo.

```{r}
# this is having issues
getmod_re <- function(df) {
  glmmTMB(y ~ year0 + (1 + year0 | cohort10t),
          data = df)
}

# lm version 1 
getmod_lm_b <- function(df) {
  lm(y ~ cohort10t,
     data = df)
}

getmod_lm_bw <- function(df) {
  lm(y ~ year0 * cohort10t,
     data = df)
}

# lm version 2 ********************************** ALLOWS ONE BEND *******************************
library(splines)
getmod_lm_b2 <- function(df) {
  lm(y ~ cohort10t,
     data = df)
}

getmod_lm_bw2 <- function(df) {
  lm(y ~ ns(year0, df = 2) * cohort10t,
     data = df)
}

# get R2
get_r2 <- function(mod) {
  r2 <- r2(mod) %>% .[[1]] %>% as.numeric()
  return(r2)
}

```

We are now ready to nest the df.

```{r}
d30_nest <- d30 %>% 
  group_by(country, variable) %>% 
  nest()
```

Estimate models for each combo, tally up the variance. (NOTE: might be using version 2 of function above.)

```{r}
results <- d30_nest %>% 
  mutate(mb = map(data, getmod_lm_b2),
         mbw = map(data, getmod_lm_bw2),
         rb = map_dbl(mb, get_r2),
         rbw = map_dbl(mbw, get_r2),
         pbetween = rb/rbw) %>% 
  select(country, variable, rb, rbw, pbetween)

ggplot(results,
       aes(x = rbw,
           y = pbetween,
           color = variable,
           label = country)) +
  geom_point() +
  geom_text_repel()

results %>% 
  ungroup() %>% 
  group_by(country) %>% 
  summarize(mb = mean(pbetween))

```



TEST NEW

```{r}
dt <- d30 %>%
  filter(country == "ZAF",
         variable == "just_prostitution") %>% 
  mutate(c10tf = factor(cohort10t))

mt <- lm(y ~ year_survey * cohort10t,
         data = dt)

p0 <- dt %>%
  group_by(c10tf, year_survey) %>%
  summarize(y = mean(y)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year_survey,
             y = y,
             group = c10tf,
             color = c10tf)) +
  geom_line()

pquad <- ggplot(dt,
             aes(x = year_survey,
                 y = y,
                 group = c10tf,
                 color = c10tf)) +
  geom_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              se = FALSE)


p1 <- ggplot(dt,
             aes(x = year_survey,
                 y = y,
                 group = c10tf,
                 color = c10tf)) +
  geom_smooth(method = "lm",
              se = FALSE)

p2 <- ggplot(dt,
             aes(x = year_survey,
                 y = y,
                 group = c10tf,
                 color = c10tf)) +
  geom_smooth(method = "lm",
              formula = y ~ 1,
              se = FALSE)

(p0 + pquad) / (p1 + p2)


```

